# Data transformation

```{r, message=FALSE}
# load packages
library(tidyverse)
library(stringr)
```


## Data preparation for high school directory data and demographics join
```{r, message=FALSE, include= FALSE}
# read in the datasets
hs_demo <- read_csv("datasets/hs_demo.csv");

hs_dir.2020 <- read_csv("datasets/2020_doe_hs_dir.csv");

hs_dir.2019 <- read_csv("datasets/2019_doe_hs_dir.csv")
```


### Demographics data

Grades included in the demographics data
```{r}
names(hs_demo)[grepl("Grade", names(hs_demo))];
```

School years included in the demographics data, and some variables included in the demographics data.
```{r}
unique(hs_demo["Year"]); names(hs_demo)[grepl("%", names(hs_demo))]; head(names(hs_demo))
```

We can see that the data contains data not only for high school, and for school years other than 2019.

The column names are difficult to retract due to spaces and special characters, so we rename columns that is easier to retract.

* replace "%" to "pct" and spaces to "_"
```{r}
# rename names easy to retract
names(hs_demo) <- str_replace_all(names(hs_demo), c(" "="_",  "%"="pct", "#"="n"))
names(hs_demo)[grepl("pct", names(hs_demo))]; head(names(hs_demo))
```


We don't want schools that are not high schools; the grade range is different per school, so we filter out schools that do not have grades 10, 11, and 12. 
Then, create new columns that contain population counts for high school grades and non-highschool grades.

Here, we set for high school to be from grades 9 to 12, according to [New York city Department of Education description](https://www.schools.nyc.gov/enrollment/enroll-grade-by-grade/high-school). It is given that students in their 8th grade apply to high schools. We assume that schools that have grades 6 to 12, or K to 12, etc. have continuing students.

Also, the dataset consists of information ranging from school years 2015 to 2019. We want to filter out years other than 2019 for initial processing of the data since the directory data are for 2019-2020.

Moreover, we created new columns `hs_enroll` and `non_hs` to divide population that are in grades 9 to 12 to others.

```{r}
# we dont want schools that are not high schools
# create a new column that would contain enrollment for only those from Grade 9(hs) and other

hs_demo_tidy <- hs_demo %>% 
  mutate(hs_enroll = Grade_9 + Grade_10 + Grade_11 + Grade_12,
         non_hs = rowSums(hs_demo[6:14]),
         PK = `Grade_3K+PK_(Half_Day_&_Full_Day)`) %>%
    dplyr::filter(Grade_9 != 0 & Year == "2019-20") %>% 
  dplyr::select(DBN, School_Name, Year, Total_Enrollment, hs_enroll, non_hs, everything(), -`Grade_3K+PK_(Half_Day_&_Full_Day)`) %>%
  dplyr::filter( Grade_10 !=0 & Grade_11 !=0 & Grade_12 != 0 )

head(hs_demo_tidy)
```


### High school directory data

Dimension of the directory data:
```{r}
dim(hs_dir.2019)
```

The directory data has very extensive information each schools, from basic school information to detailed descriptions of each program and admission information.
However, we do not want more than 400 variables, so we select only the information relevant and informative to us.
Such variables are: basic school information(DBN, school name, neighborhood) to characteristic information(programs, statistics, admissions).

The directory data has very extensive information each schools, from basic school information to detailed descriptions of each program and admission information.
However, we do not want more than 400 variables, so we select only the information relevant and informative to us.
Such variables are: basic school information(DBN, school name, neighborhood) to characteristic information(programs, statistics, admissions).


```{r}
hs_dir_tidy.2019 <- hs_dir.2019 %>% 
    dplyr::select(dbn, school_name, Borough, neighborhood, total_students, grades2018, finalgrades, program1, program2,program3,
                program4, program5, program6, program7, program8, program9, program10,
                language_classes, advancedplacement_courses, diplomaendorsements,graduation_rate,
                attendance_rate, pct_stu_enough_variety, college_career_rate, specialized, 
                admissionspriority11, admissionspriority21)

head(hs_dir_tidy.2019)
```


### Join data

Since we have the names and the DBN which are unique, we join the two datasets using these columns as indicators.

```{r}
hs_join.2019 <- hs_dir_tidy.2019 %>% 
  left_join(hs_demo_tidy, by = c("dbn" = "DBN"))%>% 
  dplyr::select(dbn, school_name, neighborhood, total_students, Total_Enrollment, hs_enroll, non_hs, grades2018, PK,
                Grade_K:Grade_12, everything()) %>% 
  mutate(across(where(is.character), ~na_if(., "N/A"))) %>% 
  column_to_rownames(var = "school_name")

print(head(hs_join.2019))
```

It is interesting that the total number of students are not equal in both data sets.

Number of observations with different reported population:
```{r}
mismatch.n.19 <- hs_join.2019[which(hs_join.2019$total_students != hs_join.2019$Total_Enrollment), ]; dim(mismatch.n.19)
```

Number of observations with identical reported population:
```{r}
match.n.19 <- hs_join.2019[which(hs_join.2019$total_students == hs_join.2019$Total_Enrollment), ]; dim(match.n.19)
```

Only 5 schools have total number of students identical in both data sets.


Investigating the data, we see that there are population data missing for 2 schools from the `Demographics` dataset.
Column `Total_Enrollment` comes from Demographics data, and `total_students` come from Directory data.

Schools that are in the Directory data but not in the demographics data:
```{r}
hs_join.2019[which(is.na(hs_join.2019["Total_Enrollment"])), 3:4]
```

```{r}
hs_join.2019[which(is.na(hs_join.2019["Total_Enrollment"])), c('grades2018', 'finalgrades')]
```

We can see that these two schools share that they do not serve grades until 12th grade.
According to the aforementioned NYC high school description (grade 9 to 12), we can filter out these schools.
Also, since the demographic data are from the Demographics dataset, we recalculate the other values into proportion with the demographics total population.

```{r}
hs_join.2019 <- hs_join.2019 %>% 
  filter(!is.na(Total_Enrollment)) %>% 
  mutate(attendance_rate = attendance_rate * total_students/Total_Enrollment,
        graduation_rate = as.numeric(graduation_rate)* total_students/Total_Enrollment ,
        pct_stu_enough_variety = pct_stu_enough_variety*total_students/Total_Enrollment ,
        college_career_rate = as.numeric(college_career_rate) * total_students/Total_Enrollment)
```

Take out the `total_students` column and the `School_Name` column, since two dataframes were joined by `dbn` and the school names are written differently.

```{r}
hs_final <- hs_join.2019 %>% 
  dplyr::select(-c(School_Name, Total_Enrollment))
```



### High school location data
```{r}
# location dataset for mapping coordinates
hs_location <- hs_dir.2019 %>% 
  dplyr::select(dbn,school_name,  Borough, city, zip, Latitude, Longitude, `Census Tract`, NTA)

head(hs_location)
```

This dataframe contains only the information on the location extracted from the high school directory file.